---
title: "R Group Project-Cleaning data"
author: "kangwei"
date: "October 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraryLoad, include=FALSE}
library(lubridate)
library(dplyr)
library(knitr)
library(stringr)
```

## Stock datasets

The dataset selected is a stock datasets. It contain details about some stocks that have been collected over several years, 2012 to 2017. The dataset for the stocks are stored on 20 CSV files. There is another CSV file namely securities.csv that contains the information about the stock companies.

First the 20 csv files are loaded at once and rbind the data frames together, since they have the same columns.

However, not all the columns names in the files are identical, different files have some differences in spacing. The data frames can't be rbind if the column names are not identical. Therefore, the column names are given in a vector rather than using what is given on the CSV files. With the same column names for all the dataframes, then it can be rbind into a single dataframe.

### Load the dataset
```{r loadData}
setwd('./nyse-financial-stocks/')
files <- list.files(pattern = "*.csv")
# The column names for the imported stock datasets
stockCol <- c("Date", "Stock", "Location", "High", "Low", "Open", "Close", 
              "Aggregated data 2 days", "Aggregated data 3 days", "Aggregated data 5 days",
              "Volume", "Number of employees", "Net change 0-numeric", "Net change 0-nominal",
              "Net change 5-numeric", "Net change 5-nominal", "Net change 25-numeric", 
              "Net change 25-nominal")
# Read the 20 source files and merge them into a single dataframe
stockDf <- do.call(rbind, lapply(files, function(x) read.csv(x, stringsAsFactors = FALSE,
                                                             header = TRUE, col.names = stockCol)))

# The column names for the imported company dataset
companyCol <- c('Stock', 'Company', 'SEC filings', 'GICS Sector', 'GICS Sub Industry', 'HQ Address',
                'Date first added', 'CIK')
companyDetails <- read.csv('../securities.csv', stringsAsFactors = FALSE, col.names = companyCol)
```

Dimension of the dataset loaded
```{r dimension}
dim(stockDf)
dim(companyDetails)
```

<br>

Structure of the *stock* data frame
```{r stockStructure}
str(stockDf)
```
Date column is not in date format.<br>
Number of employees column is in character, not numeric.

<br>

Structure of the *company details* data frame
```{r companyStructure}
str(companyDetails)
```
Date first added is not in date format.<br>

<br>

### Fix the format of the data
Fix the date formats in the dataset
```{r dateFormatting}
stockDf$Date <- dmy(stockDf$Date)

companyDetails$Date.first.added <- dmy(companyDetails$Date.first.added)
```

Fix the number of employees format into numeric
```{r changeToNumeric}
stockDf$Number.of.employees <- str_replace(stockDf$Number.of.employees, pattern = ',', replacement = '')
stockDf$Number.of.employees <- as.numeric(stockDf$Number.of.employees)
```

Some of the text would be better if formatted into factors
```{r textToFactors}
stockDf$Net.change.0.nominal <- as.factor(stockDf$Net.change.0.nominal)
stockDf$Net.change.5.nominal <- as.factor(stockDf$Net.change.5.nominal)
stockDf$Net.change.25.nominal <- as.factor(stockDf$Net.change.25.nominal)

companyDetails$SEC.filings <- as.factor(companyDetails$SEC.filings)
companyDetails$GICS.Sector <- as.factor(companyDetails$GICS.Sector)
companyDetails$GICS.Sub.Industry <- as.factor(companyDetails$GICS.Sub.Industry)

levels(stockDf$Net.change.0.nominal) <- c("Unknown", "Equal", "Negative", "Positive")
levels(stockDf$Net.change.5.nominal) <- c("Unknown", "Equal", "Negative", "Positive")
levels(stockDf$Net.change.25.nominal) <- c("Unknown", "Equal", "Negative", "Positive")

levels(stockDf$Net.change.0.nominal)
levels(stockDf$Net.change.5.nominal)
levels(stockDf$Net.change.25.nominal)

levels(companyDetails$SEC.filings)
levels(companyDetails$GICS.Sector)
levels(companyDetails$GICS.Sub.Industry)
```
<br>

Check the structure of the data after converting
```{r formatVerification}
str(stockDf)
str(companyDetails)
```
The incorrect data formats are now correctly represented.

<br>

### View the data
```{r view}
kable(head(stockDf, 10), format="markdown")
kable(head(companyDetails, 10, format='markdown'))
```

<br>

### NAs management
Check if the data has NA
```{r checkNa}
sum(is.na(stockDf))
sum(is.na(companyDetails))
```

Cleansing the NAs of the stock dataframe
``` {r naCleansingStock}
# stockDf[!complete.cases(stockDf),]
nrow(stockDf[!complete.cases(stockDf),])
# Most of the NAs are from the column Net.change.25.numeric
sum(is.na(stockDf$Net.change.25.numeric))
# Fill in the NAs with the mean from the column
stockDf$Net.change.25.numeric[is.na(stockDf$Net.change.25.numeric)] <- mean(stockDf$Net.change.25.numeric, 
                                                                            na.rm=T)
# Another column with NAs
sum(is.na(stockDf$Net.change.5.numeric))
stockDf$Net.change.5.numeric[is.na(stockDf$Net.change.5.numeric)] <- mean(stockDf$Net.change.5.numeric, 
                                                                            na.rm=T)
```

<br>

Cleansing the NAs of the company dataframe
```{r naCleasingCompany}
# Find the NAs in the company dataframe
# companyDetails[!complete.cases(companyDetails),]
nrow(companyDetails[!complete.cases(companyDetails),])
sum(is.na(companyDetails$Date.first.added))

# Replace the NAs with the date column with a date
companyDetails$Date.first.added[is.na(companyDetails$Date.first.added)] <- dmy('01-01-1970')

```

Check if all the NAs are removed
```{r naCheck}
sum(is.na(stockDf))
sum(is.na(companyDetails))
```
<br>

### Pre-analysis
Check the number of distinct stocks in the dataset
```{r checkDistinctStock}
unique(stockDf$Stock)
length(unique(stockDf$Stock))

unique(companyDetails$Stock)
length(unique(companyDetails$Stock))
```

<br>

Functions to get stock details
```{r stockDetails}
# Function  to get stock details
getCompanyDetails <- function(tickerSym) {
  stock <- companyDetails[which(companyDetails$Stock==tickerSym),]
  return(stock)
}

getCompanyName <- function(tickerSym) {
  companyName <- companyDetails[which(companyDetails$Stock==tickerSym),]$Company
  return(companyName)
}

getCompanyDetails(tickerSym = 'AXP')
getCompanyDetails(tickerSym = 'BAC')
getCompanyDetails('C')

getCompanyName(tickerSym = 'AXP')
getCompanyName(tickerSym = 'ABT')
getCompanyName(tickerSym = 'ALB')
```

<br>

### Merge the dataset
```{r mergeDataset}
stockData <- left_join(stockDf, companyDetails, by = c('Stock'))
colnames(stockData)
```

<br>

### Remove some columns
```{r removeCol}
stockData <- select(stockData, c("Date", "Stock", "High", "Low", "Open", "Close", 
                                 "Volume", "Company", "GICS.Sector", "GICS.Sub.Industry", 
                                 "Net.change.0.numeric", "Net.change.0.nominal"))
# Rename the columns
names(stockData) <- c("Date", "StockCode", "High", "Low", "Open", "Close", "Volume", "Company", 
                      "Sector", "SubIndustry", "NetChange", "NetChangeNominal")
colnames(stockData)
```

<br>

### View the finalize dataset
```{r viewFinalizeData}
kable(head(stockData, 10), format="markdown")
```
<br>

### Function to visualize the data
Function to plot the closing price
```{r plotCloseFunction}
plotClosing <- function(code) {
  subStock <- stockData %>% filter(StockCode==code, Date >= dmy('01-01-2017'))
  plot(x = subStock$Date, y = subStock$Close, main=paste("Closing price for", getCompanyName(code)),
       xlab="Year 2017", ylab="Closing Price, $", type='l')
}

plotClosing(code="BAC")
```

Function to plot the High-Low difference
```{r plotHighLowFunction}
plotHighLowDiff <- function(code) {
  subStock <- stockData %>% filter(StockCode==code, Date >= dmy('01-01-2017'))
  plot(x = subStock$Date, y = subStock$HighLowDiff, main = paste("High-Low Difference for", getCompanyName(code)), xlab = 'Year 2017', ylab = 'High-Low Difference, $', type = 'l')
}

plotHighLowDiff(code='BAC')
```

<br>